---
title: "Bigmart Sales"
author: "Yatharth Malik"
date: "January 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,message=F,warning=F}
library(caret)
library(plyr)
library(dplyr)
library(dummies)
library(mlr)
library(rpart)
library(rpart.plot)
library(caret)
library(e1071)
library(Metrics)
library(randomForest)
```

## Loading data and exploration
```{r}
train = read.csv("train.csv",na.strings = c(""," ",NA,"NA"))
test = read.csv("test.csv",na.strings = c(""," ",NA,"NA"))

summary(train)
str(train)
summary(test)
str(test)
```


Infrences drawn from data exploration :-

1. Factor mismatch in Item_Fat_Content.

2. Missing values in  Item_Weight and Outlet_Size.

3. Minimum value of Item_Visibility is 0,which is not practically possible.Hence,we'll deal them as missing values.

## Univariate Analysis
```{r}
boxplot(train$Item_MRP,main = "Boxplot of Item MRP")
boxplot(train$Item_Weight,main = "Boxplot of Item Weight")
boxplot(train$Item_Visibility,main = "Boxplot of Item Visibility")
```

## Bivariate Analysis
```{r}
ggplot(train,aes(x=Item_Visibility,y=Item_Outlet_Sales)) + geom_point(color = "red") + ggtitle("Item Visibility vs Item Outlet Sales")

ggplot(train,aes(x=Outlet_Identifier,y= Item_Outlet_Sales)) + geom_bar(stat="identity",color = "blue") + ggtitle("Outlet vs Outlet Sales") + theme(axis.text.x = element_text(angle = 70,vjust = 0.5,color = "blue"))

ggplot(train,aes(x=Item_Type,y= Item_Outlet_Sales)) + geom_bar(stat="identity",color = "purple") + ggtitle("Item Type vs Item Sales") + theme(axis.text.x = element_text(angle = 70,vjust = 0.5,color = "black"))


```
## Dealing with categorical and continuous variables

We will use median imputation to deal with continuous missing values
```{r}
test$Item_Outlet_Sales = 1
comb = rbind(train,test)

comb$Item_Weight[is.na(comb$Item_Weight)] = median(comb$Item_Weight,na.rm = T)

comb$Item_Visibility = ifelse(comb$Item_Visibility==0,median(comb$Item_Visibility),comb$Item_Visibility)

comb$Outlet_Size = ifelse(is.na(comb$Outlet_Size),"Others",comb$Outlet_Size)
comb$Outlet_Size = as.factor(comb$Outlet_Size)
levels(comb$Outlet_Size)[1] = "High"
levels(comb$Outlet_Size)[2] = "Medium"
levels(comb$Outlet_Size)[3] = "Low"

table(comb$Item_Fat_Content)

comb$Item_Fat_Content = revalue(comb$Item_Fat_Content,c("LF" = "Low Fat","reg"="Regular"))
comb$Item_Fat_Content = revalue(comb$Item_Fat_Content,c("low fat"="Low Fat"))

table(comb$Item_Fat_Content)


```

## Feature Engineering
```{r}
temp = comb%>%group_by(Outlet_Identifier)%>%tally()
names(temp)[2]  = "Outlet_Count"
comb = full_join(comb,temp,by = "Outlet_Identifier")

temp1 = comb%>%group_by(Item_Identifier)%>%tally()
names(temp1)[2]  =  "Item_Count"
comb = merge(comb,temp1,by = "Item_Identifier")

temp2 = comb%>%select(Outlet_Establishment_Year)%>%mutate(Outlet_Year = 2013 - comb$Outlet_Establishment_Year)
temp2$Outlet_Establishment_Year = NULL
comb = cbind(comb,temp2 )


items = substr(comb$Item_Identifier,1,2)
items = gsub("FD","Food",items)
items = gsub("DR","Drinks",items)
items = gsub("NC","Non Consumable",items)
comb$Item_Type_New = factor(items)

str(comb)
```

## One Hot Encoding
```{r}
comb = dummy.data.frame(comb,names = c("Outlet_Size","Outlet_Location_Type","Outlet_Type","Item_Type_New","Item_Fat_Content"),sep='_')
str(comb)
```

## Predictive Modelling

```{r}
comb = select(comb,-c(Item_Identifier,Outlet_Identifier,Item_Type,Outlet_Establishment_Year))

new_train = comb[1:nrow(train),]
new_test = comb[-(1:nrow(train)),]
names(new_train) = make.names(names(new_train))
names(new_test) = make.names(names(new_test))
```
### 1. Linear Regression
```{r}
linear_model = lm(Item_Outlet_Sales ~ . ,data = new_train)
summary(linear_model)
pred_lm = predict(linear_model,type = "response")
rmse(new_train$Item_Outlet_Sales,pred_lm)
```

### 2. Decision Trees
```{r}
tree_model = rpart(Item_Outlet_Sales ~ . ,data = new_test)
summary(tree_model)
prp(tree_model)

pred_tree = predict(tree_model,type= "vector")
rmse(new_train$Item_Outlet_Sales,pred_tree)
```

### 3. Random Forest
```{r}
rf_model = randomForest(Item_Outlet_Sales ~ . ,data = new_train,mtry = 2 ,ntree = 1000)
summary(rf_model)
varImpPlot(rf_model)
pred_rf = predict(rf_model,type="response")
rmse(new_train$Item_Outlet_Sales,pred_rf)

```
